\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-17 redaction/dhp_sept2016.tex
changement encore

If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146.} By contrast, our estimation strategy allows for an explicit decomposition of the additive cost component over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension). By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. As well, we are able to do so for the multiplicative component of trade costs. We also agglomerate the two components in an unified measure of transport costs from which we exclude the composition effects. This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}). Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.

@@

Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy. Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components. \todo{Pertinence de cet argument? A reformuler.}

>>>>

If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs, by identifying the trend patterns of the ``pure'' transport costs'' (composition effects excluded) for the two additive and ad-valorem dimensions. Our empirical analysis indeed allows for a flexible estimation of the share between the additive and the ad-valorem components of transport costs, over time/product/country partner.\footnote{The empirical specification of \cite{hummels2007}, as in \cite{hummels_skiba}, assumes a constant share of additive transport costs over time, product and country partner. In \cite{hummels_skiba}, this shows up through the fact that the elasticity of freight costs to prices $\beta$ is estimated as a constant (with $0<\beta<1$, the lower $\beta$ the larger the additive cost dimension). By contrast, our estimation strategy allows for a varying $\beta$ across time, country partner and product.} Accordingly, our estimation strategy allows for an explicit decomposition of the additive cost component taking these three dimensions into account. By contrast, in its decomposition exercise to characterize the time trend in transport costs, \citet{hummels2007} does not take into account the changes in their additive dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146. } By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. As well, we are able to do so for the multiplicative component of trade costs. We also agglomerate the two components in an unified measure of transport costs from which we exclude the composition effects. This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}). Last, for sake of comparison, we do the same exercise as \citet{hummels2007}.

Our second difference with \citet{hummels2007} relates to the way we characterize the evolution of the ``pure'' transport costs over time. \citet{hummels2007} captures the trend evolution of by the yearly dummy. Rather, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample). This notably enables us to obtain measures of transport costs that are easily comparable between transport modes and transport costs components.

@@
